{\bf Intrusion Detection using Parzen-Windows on Provenance Graph Statistics}

### 2013-12-10 writeup/paper-final.tex
% of intrusions

\begin{figure}
  \label{graph-simple}
  \centering
    \includegraphics[width=0.5\textwidth]{img/graph-simple.png}
    \caption{Example of a provenance graph, showing the types of nodes and edges common in provenance graphs} 
\end{figure}

@@

We believe that the reason we had more promising results on our simulated intrusion data sets from PASS was because the provenance graphs from these workloads were much smaller (on the order of 5000 nodes). As a result of the much smaller size, small changes in provenance graph structure from an intrusion has a much profounder impact on the smaller provenance graphs. Thus, intrusions on our PASS workload had a much higher impact on the centrality calculations and resulted in more anomalous behavior in our Parzen-Window model. In the DARPA data set, we faced much larger provenance graphs (on the order of 20,000 nodes), and because of the much larger size, the effect of small changes in provenance graph structure from an intrusion had a much smaller impact on the overall structure of the graph. Intrusions did not change the centrality metrics very much in large graphs; therefore anomalous behavior went undetected, regardless of the window size we chose. Since our system could not distinguish between anomalous and normal behavior, then the true positive rate and false positive rate are essentially identical. In addition, we found that many more nodes were being flagged as intrusions (indeed, a positive result meant that a majority of nodes in a given time range were considered anomalous). This result differed greatly from our preliminary exploration, where normal nodes were consistently identified as normal. We believe this occurred because the DARPA dataset came from a real system running for days, so the variety of nodes increased, and the distribution of these metrics became more full, so complete outliers were less present. From these results, we propose the following requirements for intrusion detection models that use provenance graph statistics in order to be useful for large provenance graphs:

>>>>

\begin{figure}
  \label{graph-simple}
  \centering
    \includegraphics[width=0.5\textwidth]{img/graph-simple.png}
    \caption{Example of a simple provenance graph, showing the types of nodes and edges common in provenance graphs} 
\end{figure}

@@

We believe that the reason we had more promising results on our simulated intrusion data sets from PASS was because the provenance graphs from these workloads were much smaller (on the order of 5000 nodes). As a result of the much smaller size, small changes in provenance graph structure from an intrusion has a much profounder impact on the smaller provenance graphs. Thus, intrusions on our PASS workload had a much higher impact on the centrality calculations and resulted in more anomalous behavior in our Parzen-Window model. In the DARPA data set, we faced much larger provenance graphs (on the order of 20,000 nodes), and because of the much larger size, the effect of small changes in provenance graph structure from an intrusion had a much smaller impact on the overall structure of the graph. Intrusions did not change the centrality metrics very much in large graphs; therefore anomalous behavior went undetected, regardless of the window size we chose. Since our system could not distinguish between anomalous and normal behavior, then the true positive rate and false positive rate are essentially identical. In addition, we found that many more nodes were being flagged as intrusions (indeed, a positive result meant that a majority of nodes in a given time range were considered anomalous). In Figure 7, we see that the percentage of nodes during exploit windows were identified similarly to those during windows containing no exploit. As expected, as PSI increases, so too does the number of events marked as exploits. This result differed greatly from our preliminary exploration, where normal nodes were consistently identified as normal. We believe this occurred because the DARPA dataset came from a real system running for days, so the variety of nodes increased, and the distribution of these metrics became more full, so complete outliers were less present. 
\begin{figure}
  \label{intrusion-v-psi}
  \centering
    \includegraphics[width=0.5\textwidth]{img/intrusion_v_psi.png}
    \caption{The percentage of nodes flagged as an intrusion for various values of PSI. This figure was generated from the Week 5 Monday data, using a window size of 10 seconds. Other window sizes showed similar results. Recall that our system flags events with a percentage of 0.5 or greater as ``anomalous''} 
\end{figure}
From these results, we propose the following requirements for intrusion detection models that use provenance graph statistics in order to be useful for large provenance graphs:
